---
title: "p8105_hw5_akg2197"
author: "Annie Gorman"
date: "2024-11-12"
output: github_document
---

## Loading libraries 

```{r, include = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


# Problem 2

### Function and map
```{r}
sim_function <- function(n = 30, mu = 0, sigma = 5) {
  sim_data = tibble(
    x = rnorm(n, mean = mu, sd = sigma)
  )
  test = t.test(x ~ 1, data = sim_data)
  broom::tidy(test)
}
```

```{r}
sim_results = tibble(mu = 0:6) |>
  mutate(
    output_df = map(.x = mu, ~rerun(5000, sim_function(n = 30, mu=.x))) ,
    value_df = map(output_df, bind_rows)
  )
```

```{r}
sim_results = sim_results |>
  unnest(value_df) |>
  select(estimate, p.value, mu) |>
  mutate(significant_results = as.numeric(p.value < 0.05))
```

### Let's make some plots using the above function outputs

#### Plot showing the proportion of times the null was rejected (the power of the test) on the y axis and the true value of mu on the x axis:

```{r}
sim_results |>
  group_by(mu) |>
  summarize(power = mean(significant_results)) |>
  ggplot(aes(x = mu, y = power)) + 
  geom_path() 
```
From this plot, we can see that as effect size increases, so does power. Although this relationship is not linear, we do observe a direct relationship between effect size and power. Effect size describes the difference between the true statistics and the null statistics.

#### Plot showing the average estimate of mu on the y axis and the true value of mu on the x axis

```{r}
sim_results |>
  group_by(mu) |>
  summarize(estimatedtrue_mu = mean(estimate)) |>
  ggplot(aes(x = mu, y = estimatedtrue_mu)) + 
  geom_point() + 
  geom_path() 
```
#### Average estimate of mu only in samples for which the null was rejected on the y axis and the true value of mu on the x axis
```{r}
sim_results |>
  filter(significant_results == 1) |>
  group_by(mu) |>
  summarize(estimatedtrue_mu = mean(estimate)) |>
  ggplot(aes(x = mu, y = estimatedtrue_mu)) + 
  geom_point() + 
  geom_path() 
```

When mu is equal to 1, 2, or 3, the the sample average of mu across tests for which the null is rejected is NOT equal to the true value of mu. For mu = 4, 5, or 6, the average of mu is approximately equal to the true value of mu. This is because higher values lead to rejection of the mu (significant result). However, when we have smaller values like 1, 2, or 3, a smaller proportion of the sample gets rejected, and many of those values will have sample values larger than the true mu. This would result in unequal sample averages to the true mu, especially sample averages that are higher than the true mu.  